Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets

### 2018-08-19 doc/Response.Rmd

The article "Adapting the Chumbley Score to match striae on Land Engraved Areass (LEAs) of bullets" extends an existing comparison algorithm for striated marks to the analysis of bullet LEAs. The paper makes two primary contributions. First, they modify the deterministic work of Hadler & Morris (2017) which itself is an extension of Chumbley et al. The authors' contribution is a modification of Hadler & Morris to reduce the number of scenarios which 'fail' under the Hadler & Morris approach. Second, they evaluate the work on two sets of bullet LEA data and explore some variants of the algorithm parameters to tune it for bullets rather than screwdrivers. Overall the paper presents a reasonable project to extend a nice algorithm for striated toolmark comparison.

Adaptation of the Chumbley el al algorithm from screwdrivers to bullets makes sense as both analyses involve striated marks. The authors here nicely point out that compared to screwdriver marks (as in the original Chumbley paper), LEAs are typically smaller both in length and width. LEAs also have a curved cross-section baseline which must be corrected.

The authors do a nice job covering previous work. I appreciate the summary tables 1 and 2. Table 1 is good as-is. Table 2 can be a bit misleading because we don't know what bullet datasets were used. Because these different papers analyzed different bullet data sets the underlying difficulty of matching impacts the performance. That is, bullets from very well marking firearms are 'easier' to match than those from firearms which mark poorly. So the FP and FN reported should be taken with a grain of salt. The authors should add a statement regarding different performance on different datasets.

*Thank you for pointing this out. You are right, we have added more information about the data, wherever available. We have also added a sentence to emphasize different difficulty levels and resulting differences in error rates of different brands of firearms and ammunition*

### 2018-08-19 doc/paper.Rmd
resolution of images

---
title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
affiliation: Department of Statistics, Iowa State University and CSAFE (Center for
  Statistics and Applications in Forensic Evidence)
authors:
- degree: Master of Science (M.Sc.)
  name: Ganesh Krishnan,
- degree: Doctor of Philosophy (Ph.D.)
  name: Prof. Heike Hofmann,
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
bibliography: bibliography
blinded: 0
funding: The authors gratefully acknowledge funding from CSAFE through cooperative
  agreement \# 70NANB15H176 between NIST and Iowa State University.
keywords:
- forensic science
- toolmark
- cross-correlation
- Mann-Whitney U statistic
- land engraved areas (LEAs)
- algorithm
- signatures
- same-source problem
biblio-style: jfs
presentedat: This work was presented in the form of a poster in the AFTE 2018 49th
  Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West
  Virginia
thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab
  in scanning the Hamby-44 set and providing the scans to us. We would also like to
  thank NIST for providing the Hamby-44 bullets to us for scanning.
abstract: The same-source problem remains a major challenge in forensic toolmark and
  firearm examination. Technological advances in surface imaging allow measurements
  of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
  Here, we investigate the applicability of the Chumbley scoring method \cite{hadler},
  developed for screwdriver markings, for same-source identification of bullet striae.
  We provide methods to identify parameters that minimize the error rates for matching
  of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE. We suggest a
  remedial algorithm to alleviate the problem of failed tests in the method, increasing
  both the power of the test and reducing error rates. Type II error rates of the
  proposed method improve by more than one third (Type I error of 0.05) and are on
  average about 0.22. This puts the proposed method on similar footing as other single
  feature matching approaches in the literature.
---

@@

```{r, fig.cap="Overview of all of the results."}
rocs <- roc.summary %>% dplyr::select(-auc, -coarse, -alpha)
rocs <- rocs %>% mutate(
  aucplus = sprintf("%.3f (%.3f, %.3f)", AUC, lower, upper)
)
rocs <- rocs %>% dplyr::filter(wv==30)
rocs <- rocs %>% dplyr::select(-AUC, -lower, -upper, -wv)
rocs1 <- rocs %>% dplyr::filter(method=="CS1")
rocs2 <- rocs %>% dplyr::filter(method=="CS2")
rocs_tab <- data.frame(rocs1 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
                  rocs2 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
dt <- xtable(rocs_tab, 
             digits=c(0,0,3, 3,0,3,3,0))
#print(dt, include.rownames = FALSE)
```

\begin{table}[hbtp]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\ 
  \hline
90 & 0.068 & 0.330 & 0.850 (0.835, 0.865) & 0.075 & 0.242 & 0.877 (0.863, 0.890) \\ 
  120 & 0.065 & 0.309 & 0.864 (0.850, 0.878) & 0.072 & 0.236 & 0.890 (0.877, 0.903) \\ 
  130 & 0.065 & 0.317 & 0.863 (0.850, 0.877) & 0.072 & 0.224 & 0.898 (0.886, 0.910) \\ 
  140 & 0.065 & 0.304 & 0.873 (0.859, 0.886) & 0.071 & 0.229 & 0.902 (0.891, 0.914) \\ 
  150 & 0.067 & 0.326 & 0.863 (0.850, 0.877) & 0.072 & 0.225 & 0.904 (0.892, 0.916) \\ 
  180 & 0.061 & 0.326 & 0.877 (0.864, 0.890) & 0.070 & 0.229 & 0.907 (0.896, 0.919) \\ 
  210 & 0.064 & 0.342 & 0.865 (0.851, 0.878) & 0.066 & 0.231 & 0.906 (0.895, 0.918) \\ 
   \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}. FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate. Area under the curve (AUC) is shown with confidence intervals. }
\end{table}

@@

\autoref{fig:type1} shows observed type I error rates (FPR) across a range of optimization windows $w_o$. Generally, observed type I errors are higher than expected. Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I error rates.

@@

The high-resolution scans of Hamby set 44 are capturing images at a resolution of $0.645 \mu m$ per pixel. On average, land engraved areas are 3000 pixels in length. A coarseness of $c = 0.125$ seemed to be sufficient in removing any bullet curvature. Both methods have a failed test rate of less than 0.006, indicating, again, that the larger number of pixels alleviates the problem of test failures.
\autoref{fig:h44} shows the resulting EER and AUC for methods CS1 and CS2 based on two sizes of validation windows ($w_v \in 75, 125$) and optimization window sizes around 300 pixels (10 percent of the average length). Both methods show an increase in performance around $w_o = 300$ pixels. CS2 out-performs CS1 in all scenarios, but the difference is not significant (using DeLong's confidence intervals). Interestingly, the overall performance of both CS1 and CS2 is a lot lower for the high resolution version of Hamby-44 than for the lower resolution scans. The area under the curve overall is significantly lower for the high resolution scans than for the previous set of scans. Partly, this might be due to the particular choice of the parameters, partly the higher-resolution scans might be picking up on real differences between the lands that the lower-resolution scans fail to detect.

@@

```{r, fig.cap="h44 Overview of all of the results."}
colnames(h44)[11]<- "coarse"
roc44.summary <- h44 %>% group_by(wo, wv, method, coarse) %>%
  summarize(
    eer = EER(1-p_value, match),
    auc = list(pROC::ci.auc(match, 1-p_value))
  )

@@

rocs44 <- roc44.summary %>% dplyr::select(-auc, -coarse, -alpha)
rocs44 <- rocs44 %>% mutate(
  aucplus = sprintf("%.3f (%.3f, %.3f)", AUC, lower, upper)
)
rocs44 <- rocs44 %>% dplyr::filter(wv==75, wo %% 30 == 0)
rocs44 <- rocs44 %>% dplyr::select(-AUC, -lower, -upper, -wv)
rocs1.44 <- rocs44 %>% dplyr::filter(method=="cs1")
rocs2.44 <- rocs44 %>% dplyr::filter(method=="cs2")
#rocs2.44<- rocs2.44[-3,]
rocs44_tab <- data.frame(rocs1.44 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
                       rocs2.44 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
dt44 <- xtable(rocs44_tab, 
             digits=c(0,0,3, 3,0,3,3,0))
#print(dt44, include.rownames = FALSE)
```

\begin{table}[hbtp]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) & \bf FPR & \bf FNR & \bf AUC (95\% C.I.) \\  
  \hline
%90 & 0.077 & 0.546 & 0.680 (0.633, 0.727) & 0.056 & 0.478 & 0.729 (0.684, 0.774) \\
%  120 & 0.082 & 0.485 & 0.736 (0.693, 0.779) & 0.056 & 0.452 & 0.763 (0.720, 0.806) \\
%  150 & 0.080 & 0.493 & 0.730 (0.685, 0.775) & 0.055 & 0.422 & 0.774 (0.733, 0.816) \\
%  180 & 0.087 & 0.458 & 0.773 (0.734, 0.813) & 0.055 & 0.435 & 0.773 (0.730, 0.816) \\
  210 & 0.086 & 0.413 & 0.780 (0.740, 0.821) & 0.055 & 0.387 & 0.803 (0.763, 0.843) \\ 
  240 & 0.085 & 0.427 & 0.784 (0.744, 0.824) & 0.055 & 0.365 & 0.801 (0.760, 0.841) \\ 
  270 & 0.082 & 0.415 & 0.782 (0.741, 0.823) & 0.048 & 0.374 & 0.808 (0.769, 0.848) \\ 
  300 & 0.082 & 0.405 & 0.791 (0.751, 0.832) & 0.051 & 0.383 & 0.814 (0.776, 0.852) \\ 
  330 & 0.082 & 0.428 & 0.791 (0.753, 0.830) & 0.050 & 0.383 & 0.804 (0.765, 0.844) \\ 
  360 & 0.083 & 0.427 & 0.788 (0.747, 0.828) & 0.049 & 0.400 & 0.803 (0.765, 0.842) \\ 
  390 & 0.085 & 0.400 & 0.789 (0.748, 0.830) & 0.050 & 0.383 & 0.810 (0.771, 0.849) \\ 
   \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}. FPR is the observed rate of false positives (for a fixed level $\alpha = 0.05$), FNR is the false negative rate. Area under the curve (AUC) is shown with 95\% confidence intervals. }
\end{table}

@@

In assessing the suitability of the (deterministic) Chumbley Score for matching striae on bullet lands we have gained valuable insights into the process: method CS1 as proposed by \citet{hadler} has a strong dependency on the specific choice of parameters; the defaults suggested by \citet{hadler} for screw drivers are not directly applicable  for the smaller bullet lands. The coarseness parameter in particular has a strong impact on the performance of the test. However, we were able to suggest some heuristics based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile. For bullet lands we found  a coarseness value of $c = 0.15$ to be suitable for the low-resolution scans from NIST and a value of $c=0.125$ suitable for the higher-resolution scans from CSAFE.
Sizes for optimization windows $w_o$ were based on cross-validation to minimize overall type 2 error rates. Unfortunately, this kind of assessment is only feasible in the setting of a large study, such as the one we presented, and does not transfer immediately to case work, where a forensic examiner would only deal with a few identifications. Parameter settings for different studies should be investigated and reported to allow for a further refinement.

Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of 0.272 for an optimized window size of 140 pixels -- which is considerably higher than the error rates achieved on matching toolmarks, but is similar to other single-feature methods proposed for bullet matching. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
 Algorithm CS2 is introduced here as a remedy for failed tests by introducing an alternate version of choosing different-shift pairs. 
The algorithm CS2 introduced here is constructed in a way that achieves on average a ten-fold reduction in the number of failures. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates of CS2 reach a minimum of 0.217 for an optimized window size of 130 pixels. This increase in power of CS2 over CS1 should also apply to previous studies on toolmarks. It would be interesting to see these results using the adjusted algorithm. Unfortunately, none of the studies have made the data publicly accessible.

@@

However, bullets usually have multiple lands -- in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. We might be able to get more power out of the test by adapting CS2 to work in a bullet-to-bullet comparison.

>>>>

---
title: "Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets"
affiliation: Department of Statistics, Iowa State University and CSAFE (Center for
  Statistics and Applications in Forensic Evidence)
authors:
- degree: Master of Science (M.Sc.)
  name: Ganesh Krishnan,
- degree: Doctor of Philosophy (Ph.D.)
  name: Prof. Heike Hofmann,
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
bibliography: bibliography
blinded: 0
funding: The authors gratefully acknowledge funding from CSAFE through cooperative
  agreement \# 70NANB15H176 between NIST and Iowa State University.
keywords:
- forensic science
- toolmark
- cross-correlation
- Mann-Whitney U statistic
- land engraved areas (LEAs)
- algorithm
- signatures
- same-source problem
biblio-style: jfs
presentedat: This work was presented in the form of a poster in the AFTE 2018 49th
  Annual Training Seminar, held from June 3rd to June 8th, 2018 in Charleston, West
  Virginia
thank: We would like to thank the efforts of CSAFE and Roy J Carver Microscopy lab
  in scanning the Hamby-44 set and providing the scans to us. We would also like to
  thank NIST for providing the Hamby-44 bullets to us for scanning.
  We would also like to thank the two anonymous reviewers for their helpful suggestions in improving the paper.
abstract: The same-source problem remains a major challenge in forensic toolmark and
  firearm examination. Technological advances in surface imaging allow measurements
  of 3D surfaces at previously unforeseen resolutions and enable digitized imaging.
  Here, we investigate the applicability of the Chumbley scoring method \cite{hadler},
  developed for screwdriver markings, for same-source identification of bullet striae.
  We provide methods to identify parameters that minimize the error rates for matching
  of LEAs using Hamby datasets 44 and 252 measured by NIST and CSAFE. We suggest a
  remedial algorithm to alleviate the problem of failed tests in the method, increasing
  both the power of the test and reducing error rates. Type II error rates of the
  proposed method improve by more than one third (Type I error of 0.05) and are on
  average about 0.22. This puts the proposed method on similar footing as other single
  feature matching approaches in the literature.
---

@@

```{r, fig.cap="Overview of all of the results."}
rocs <- roc.summary %>% dplyr::select(-auc, -coarse, -alpha)
rocs <- rocs %>% mutate(
  aucplus = sprintf("%.3f (%.3f, %.3f)", AUC, lower, upper)
)
rocs <- rocs %>% dplyr::filter(wv==30)
rocs <- rocs %>% dplyr::select(-AUC, -lower, -upper, -wv)
rocs1 <- rocs %>% dplyr::filter(method=="CS1")
rocs2 <- rocs %>% dplyr::filter(method=="CS2")
rocs_tab <- data.frame(rocs1 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
                  rocs2 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
rocs_tab[,c(2,3,5,6)] <- rocs_tab[,c(2,3,5,6)]*100
dt <- xtable(rocs_tab, 
             digits=c(0,0,1, 1,0,1,1,0))
#print(dt, include.rownames = FALSE)
```

\begin{table}[hbtp]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) \\ 
  \hline
90 & 6.8 & 33.0 & 0.850 (0.835, 0.865) & 7.5 & 24.2 & 0.877 (0.863, 0.890) \\ 
  120 & 6.5 & 30.9 & 0.864 (0.850, 0.878) & 7.2 & 23.6 & 0.890 (0.877, 0.903) \\ 
  130 & 6.5 & 31.7 & 0.863 (0.850, 0.877) & 7.2 & 22.4 & 0.898 (0.886, 0.910) \\ 
  140 & 6.5 & 30.4 & 0.873 (0.859, 0.886) & 7.1 & 22.9 & 0.902 (0.891, 0.914) \\ 
  150 & 6.7 & 32.6 & 0.863 (0.850, 0.877) & 7.2 & 22.5 & 0.904 (0.892, 0.916) \\ 
  180 & 6.1 & 32.6 & 0.877 (0.864, 0.890) & 7.0 & 22.9 & 0.907 (0.896, 0.919) \\ 
  210 & 6.4 & 34.2 & 0.865 (0.851, 0.878) & 6.6 & 23.1 & 0.906 (0.895, 0.918) \\    \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucs} Overview of results as shown in Figures \ref{fig:roc} and \ref{fig:eer}. \%FP is the observed percentage of false positives (for a fixed level $\alpha = 0.05$), \%FN is the percentage of false negatives. Area under the curve (AUC) is shown with confidence intervals. }
\end{table}

@@

\autoref{fig:type1} shows the percentages of observed type I errors (\%FP) across a range of optimization windows $w_o$. Generally, observed type I errors are higher than expected. Method CS1 shows in this instance slightly better performance than method CS2, but for both an increase in the size of the optimization window leads to a decrease in the observed type I errors.

@@

The high-resolution scans of Hamby set 44 are capturing images at a resolution of $0.645 \mu m$ per pixel. On average, land engraved areas are 3000 pixels in length. A coarseness of $c = 0.125$ seemed to be sufficient in removing any bullet curvature. Both methods have a failed test rate of less than 0.006, indicating, again, that the larger number of pixels alleviates the problem of test failures.
\autoref{fig:h44} shows the resulting EER and AUC for methods CS1 and CS2 based on two sizes of validation windows ($w_v \in 75, 125$) and optimization window sizes around 300 pixels (10 percent of the average length), with corresponding errors shown numerically in \autoref{tab:aucsh44}. Both methods show an increase in performance around $w_o = 300$ pixels. CS2 out-performs CS1 in all scenarios, but the difference is not significant (using DeLong's confidence intervals). Interestingly, the overall performance of both CS1 and CS2 is a lot lower for the high resolution version of Hamby-44 than for the lower resolution scans. The area under the curve overall is significantly lower for the high resolution scans than for the previous set of scans. Partly, this might be due to the particular choice of the parameters, partly the higher-resolution scans might be picking up on real differences between the lands that the lower-resolution scans fail to detect. Re-assuringly, the observed error rate of false positives for CS2 is closer to the nominal rate of 5\% than for the lower-resolution scans.

@@

```{r, fig.cap="Overview of all of the results."}
colnames(h44)[11]<- "coarse"
roc44.summary <- h44 %>% group_by(wo, wv, method, coarse) %>%
  summarize(
    eer = EER(1-p_value, match),
    auc = list(pROC::ci.auc(match, 1-p_value))
  )

@@

rocs44 <- roc44.summary %>% dplyr::select(-auc, -coarse, -alpha)
rocs44 <- rocs44 %>% mutate(
  aucplus = sprintf("%.3f (%.3f, %.3f)", AUC, lower, upper)
)
rocs44 <- rocs44 %>% dplyr::filter(wv==75, wo %% 30 == 0)
rocs44 <- rocs44 %>% dplyr::select(-AUC, -lower, -upper, -wv)
rocs1.44 <- rocs44 %>% dplyr::filter(method=="cs1")
rocs2.44 <- rocs44 %>% dplyr::filter(method=="cs2")
#rocs2.44<- rocs2.44[-3,]
rocs44_tab <- data.frame(rocs1.44 %>% ungroup() %>% dplyr::select(-wv, -method, -eer), 
                       rocs2.44 %>% ungroup() %>% dplyr::select(actual, beta, aucplus))
rocs44_tab[, c(2,3,5,6)] <- rocs44_tab[, c(2,3,5,6)]*100
dt44 <- xtable(rocs44_tab, 
             digits=c(0,0,1,1,0,1,1,0))
#print(dt44, include.rownames = FALSE)
```

\begin{table}[hbtp]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{rrrlrrl}
  \hline
  & \multicolumn{3}{l}{\bf CS1} & \multicolumn{3}{l}{\bf CS2} \\
$w_o$ & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) & \bf \%FP & \bf \%FN & \bf AUC (95\% C.I.) \\  
  \hline
  210 & 8.6 & 41.3 & 0.780 (0.740, 0.821) & 5.5 & 38.7 & 0.803 (0.763, 0.843) \\ 
  240 & 8.5 & 42.7 & 0.784 (0.744, 0.824) & 5.5 & 36.5 & 0.801 (0.760, 0.841) \\ 
  270 & 8.2 & 41.5 & 0.782 (0.741, 0.823) & 4.8 & 37.4 & 0.808 (0.769, 0.848) \\ 
  300 & 8.2 & 40.5 & 0.791 (0.751, 0.832) & 5.1 & 38.3 & 0.814 (0.776, 0.852) \\ 
  330 & 8.2 & 42.8 & 0.791 (0.753, 0.830) & 5.0 & 38.3 & 0.804 (0.765, 0.844) \\ 
  360 & 8.3 & 42.7 & 0.788 (0.747, 0.828) & 4.9 & 40.0 & 0.803 (0.765, 0.842) \\
  390 & 8.5 & 40.0 & 0.789 (0.748, 0.830) & 5.0 & 38.3 & 0.810 (0.771, 0.849) \\ 
   \hline
\end{tabular}
\end{adjustbox}
\caption{\label{tab:aucsh44} Overview of high resolution scan results as shown in Figures \ref{fig:h44}. \%FP is the percentage of observed  false positives (for a fixed level $\alpha = 0.05$), \%FN is the percentage of false negatives. Area under the curve (AUC) is shown with 95\% confidence intervals. }
\end{table}

@@

In assessing the suitability of the (deterministic) Chumbley Score for matching striae on bullet lands we have gained valuable insights into the process: method CS1 as proposed by \citet{hadler} has a strong dependency on the specific choice of parameters; the defaults suggested by \citet{hadler} for screw drivers are not directly applicable  for the smaller bullet lands. The coarseness parameter in particular has a strong impact on the performance of the test. However, we were able to suggest some heuristics based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile. For bullet lands we found  a coarseness value of $c = 0.15$ to be suitable for the low-resolution scans from NIST and a value of $c=0.125$ suitable for the higher-resolution scans from CSAFE.
Sizes for optimization windows $w_o$ were based on cross-validation to minimize overall type 2 error rates. 
Ideally, the exact values for parameters will be determined in a large study incorporating different types of firearms and brands of ammunition. 
%Unfortunately, this kind of assessment is only feasible in the setting of a large study, such as the one we presented, 
Results therefore do not transfer immediately to case work, where a forensic examiner would only deal with a few identifications. %Parameter settings for different studies should be investigated and reported to allow for a further refinement.

Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of 0.272 for an optimized window size of 140 pixels -- which is considerably higher than the error rates achieved on matching toolmarks, but is similar to other single-feature methods proposed for bullet matching. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
 Algorithm CS2 is introduced here as a remedy for failed tests by introducing an alternate version of choosing different-shift pairs. 
Algorithm CS2 is constructed in a way that achieves on average a ten-fold reduction in the number of failures. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates of CS2 reach a minimum of 0.217 for an optimized window size of 130 pixels. This increase in power of CS2 over CS1 should also apply to previous studies on toolmarks. It would be interesting to see these results using the adjusted algorithm. Unfortunately, none of the studies have made the data publicly accessible.

@@

Different avenues for improving the performance of the Chumbley-score are still open: 
(i) When making the original Chumbley Score deterministic, \citet{hadler} rely on an optimal shift based on the maximum cross-correlation between two markings. An erroneous decision in  identifying the optimial shift between markings leads almost always to a missed identification. CS2 is still susceptible to this source of errors.
(ii) Bullets usually have multiple lands -- in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. As shown by \citet{chu2010} in the example of the cross-correlation between lands, we will be able to get more power out of the test by adapting the algorithm to make use of the relative position of a land on the bullet to combine multiple land-to-land comparisons into a single bullet-to-bullet comparison.
